---
title: "STAT_167_Project"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

## Libraries
```{r}
library(tidyverse) ## For 'dplyr' & 'ggplot'
library(gridExtra)
```


```{r}
df <- read.csv("Cleaner_Data.csv", header = TRUE, sep = ',')
```

## Data Cleaning and Manipulation

```{r}
## Rename Variables
rn_df <- df %>%
  rename(Age = Age.at.enrollment,
         Mart.Stat = Marital.status,
         Mom.Occp = Mother.Occupation,
         Mom.Ed = Mother.Quals,
         Dad.Occp = Father.Occupation,
         Dad.Ed = Father.Quals,
         Smstr1.GPA = Curricular.units.1st.sem..grade.,
         Smstr2.GPA = Curricular.units.2nd.sem..grade.)

## Filter Data to a new data frame
fltr_df <- rn_df |>
  select(Target, Age, 
         Smstr1.GPA, Smstr2.GPA,
         Gender, Mart.Stat, 
         Nationality, 
         Mom.Ed, Mom.Occp,
         Dad.Ed, Dad.Occp,
         Target, Course,
         Admission.Grade,
         Scholarship,) |>
  arrange(Age) 
#fltr_df
```

```{r}
## Replace semester GPA scale from(0,20) to (0,5)
fltr_df$Smstr1.GPA <- round( (fltr_df$Smstr1.GPA/20)*5, digits = 2)
fltr_df$Smstr2.GPA <- round( (fltr_df$Smstr2.GPA/20)*5, digits = 2)

## Assign each student to an avg grade category that reflects their GPA
## UCI's Grading(http://www.reg.uci.edu/services/transcripts/notations.html)
## UCR's Grading()
asgn_Grade <- function(gpa){
  brackets = c("na")
  for(i in 1:length(gpa)){
    if(0<= gpa[i] & gpa[i] <= 0.69){
      brackets[i] <- "F"
    }else if(0.7<= gpa[i] & gpa[i] <=0.99){
      brackets[i] <- "D-"
    }else if(1<= gpa[i] & gpa[i] <=1.29){
      brackets[i] <- "D"
    }else if(1.3<= gpa[i] & gpa[i] <=1.69){
      brackets[i] <- "D+"
    }else if(1.7<= gpa[i] & gpa[i] <=1.99){
      brackets[i] <- "C-"
    }else if(2.0<= gpa[i] & gpa[i] <=2.29){
      brackets[i] <- "C"
    }else if(2.3<= gpa[i] & gpa[i] <=2.69){
      brackets[i] <- "C+"
    }else if(2.7<= gpa[i] & gpa[i] <=2.99){
      brackets[i] <- "B-"
    }else if(3.0<= gpa[i] & gpa[i] <=3.29){
      brackets[i] <- "B"
    }else if(3.3<= gpa[i] & gpa[i]<=3.69){
      brackets[i] <- "B+"
    }else if(3.7<= gpa[i] & gpa[i] <=3.99){
      brackets[i] <- "A-"
    }else if(4.0<=gpa[i]){
      brackets[i] <- "A"
    }
  }
  return(brackets)
}

## Assign each student to an age group
asgn_AgeGrp <- function(age){
  brackets <- c("na")
  for(i in 1:length(age)){
    if(17<=age[i] & age[i]<=19){
      brackets[i] <- "A_yng_adults"
    }else if(20<=age[i] & age[i]<=24){
      brackets[i] <- "B_early_20s"
    }else if(25<=age[i] & age[i]<=29){
      brackets[i] <- "C_late_20s"
    }else if(30<=age[i] & age[i]<=34){
      brackets[i] <- "D_early_30s"
    }else if(35<=age[i] & age[i]<=39){
      brackets[i] <- "E_late_30s"
    }else if(40<=age[i] & age[i]<=44){
      brackets[i] <- "F_early_40s"
    }else if(45<=age[i] & age[i]<=49){
      brackets[i] <- "G_late_40s"
    }else if(50<=age[i] & age[i]<=54){
      brackets[i] <- "H_early_50s"
    }else if(55<=age[i] & age[i]<=59){
      brackets[i] <- "I_late_50s"
    }else if(60<=age[i] & age[i]<=64){
      brackets[i] <- "J_early_60s"
    }else if(65<=age[i] & age[i]<=69){
      brackets[i] <- "K_late_60s"
    }else if(70<=age[i]){
      brackets[i] <- "L_over_70"
    }
  }
  return(brackets)
}
fltr_df$Age_grp <- asgn_AgeGrp(fltr_df$Age)

## Create new variable called GPA.ttl
fltr_df <- fltr_df |>
  mutate( GPA.Ttl = round( (Smstr1.GPA + Smstr2.GPA)/2, 2), .after = Smstr2.GPA )

#fltr_df
```


## Of those that dropped, What's their GPA info?
```{r, fig.height=5.5}
dropouts <- filter(fltr_df, Target=="Dropout")

## Graph of raw data. Too many peaks. Hard to interpret.
plot_a1 <- ggplot(data = dropouts, mapping = aes(x = (GPA.Ttl))) +
  geom_density() +
  labs(x = "GPA")

plot_a2 <- ggplot(data = dropouts, mapping = aes(x=(GPA.Ttl))) +
  geom_bar() +
  xlim(2,4) +
  labs(x = "GPA")

plot_a3 <- ggplot(data = fltr_df, mapping = aes(x = GPA.Ttl)) +
  geom_boxplot() +
  labs(x = "GPA")

plot_a4 <- ggplot(data = fltr_df, mapping = aes(x = GPA.Ttl)) +
  geom_boxplot() +
  xlim(2,4) +
  labs(x = "GPA")

grid.arrange(plot_a1, plot_a2, plot_a3, plot_a4, ncol = 2)
## Note the number of Rows being removed.
```

```{r, fig.height=5.5}
## Graph of (data)^2
plot_a5 <- ggplot(data = dropouts, mapping = aes(x = (GPA.Ttl)^2)) +
  geom_density()
## Not any better than previous graph of raw data

plot_a6 <- ggplot(data = dropouts, mapping = aes(x=(GPA.Ttl))) +
  geom_bar() +
  labs(x = "GPA")

plot_a7 <- ggplot(data = fltr_df, mapping = aes(x = (GPA.Ttl)^2)) +
  geom_boxplot() +
  labs(x = "GPA")

plot_a8 <- ggplot(data = fltr_df, mapping = aes(x = (GPA.Ttl)^2)) +
  geom_boxplot() +
  labs(x = "GPA")

grid.arrange(plot_a5, plot_a6, plot_a7, plot_a8, ncol = 2)
## Note the number of Rows being removed.
```

```{r, fig.height=5.5}
## Graph of log(data).
plot_a9 <-ggplot(data = dropouts, mapping = aes(x = log((GPA.Ttl)))) +
  geom_density()
## Few peaks, but still not quite good.

plot_a10 <- ggplot(data = dropouts, mapping = aes(x=log(GPA.Ttl))) +
  geom_bar()  +
  labs(x = "GPA")

plot_a11 <-ggplot(data = fltr_df, mapping = aes(x = log(GPA.Ttl))) +
  geom_boxplot() +
  xlim(0.75,1.5) +
  labs(x = "GPA")

plot_a12 <-ggplot(data = fltr_df, mapping = aes(x = log(GPA.Ttl))) +
  geom_boxplot() +
  labs(x = "GPA")

grid.arrange(plot_a9, plot_a10, plot_a11, plot_a12, ncol = 2)
## Note the number of Rows being removed.
```

**Interpretation From Boxplots** From the raw data boxplot and log boxplot, I see there are many outliers. Checking to see if a good portion of the dropout students' GPAs are between 2.5 and 4.0.

```{r}
## Investigating Outliers
##summary(dropouts$GPA.Ttl)

length(dropouts$GPA.Ttl) ## 1421 students
length(dropouts$GPA.Ttl) - sum(dropouts$GPA.Ttl<2.5) ## (1421 - 751) = 670 students above 2.5

## Dropouts Information
# students_Drop_info <- fltr_df |>
#   filter(Target=="Dropout") |>
#   group_by(Gender) |>
#   reframe(Total_Students = n(), 
#           Avg_GPA = mean(GPA.Ttl), 
#           Min_GPA = min(GPA.Ttl), 
#           Sum_of_MinGPA = sum(GPA.Ttl==min(GPA.Ttl)),
#           Max_GPA = max(GPA.Ttl),
#           Sum_of_MaxGPA = sum(GPA.Ttl==max(GPA.Ttl)))
# 
# students_Drop_info
```
There are more dropout students with GPA below 2.5. I can't just remove them.

## Let's look at students by their Grade to see who has the highest dropout rate

```{r, fig.height=5.5}
fltr_df$Grade <- asgn_Grade(fltr_df$GPA.Ttl)
fltr_df$Grade
fltr_df$Grade <- factor(fltr_df$Grade, levels = c("F","D-","D","D+","C-","C","C+","B-","B","B+","A-","A","A+"))

plot_b1 <- ggplot(data = fltr_df, mapping = aes(x=Grade, fill = Target)) +
  geom_bar() +
  labs(x = "Grade")

plot_b1

```

```{r}
fltr_df$Target <- factor(fltr_df$Target)
```
```{r}
# building a multiple linear regression model with binary discrete and continuous variables with the average total GPA as the response variable



# Function for backward selection using adjusted R^2


# using a backwards selection technique and removing variables based on which combination of variables result in a higher adjusted R^2 value

mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Mom.Ed + Mom.Occp + Dad.Ed + Dad.Occp + Admission.Grade + Scholarship, fltr_df)
summary(mlr_model_backwards)

#remove the variable with the highest p-value approach
mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Mom.Ed + Mom.Occp + Dad.Ed + Admission.Grade + Scholarship, fltr_df)
mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Mom.Ed + Mom.Occp + Admission.Grade + Scholarship, fltr_df)
mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Mom.Ed + Admission.Grade + Scholarship, fltr_df)
mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Admission.Grade + Scholarship, fltr_df)

#removing any more variables results in a lower adjusted R^2 value so we keep this current model
summary(mlr_model_backwards)


# repeat this process using forwards selection by starting with an empty model and chosing the variables that increase the adjusted r^2 value the most

# adjusted R^2 values for each variable

mlr_model_forwards <- lm(GPA.Ttl ~ Age, fltr_df) #0.02954

mlr_model_forwards <- lm(GPA.Ttl ~ Gender, fltr_df) #0.04093

mlr_model_forwards <- lm(GPA.Ttl ~ Mom.Ed, fltr_df) # 0.001004

mlr_model_forwards <- lm(GPA.Ttl ~ Mom.Occp, fltr_df) #0.0001845

mlr_model_forwards <- lm(GPA.Ttl ~ Dad.Ed, fltr_df) #0.0001567

mlr_model_forwards <- lm(GPA.Ttl ~ Dad.Occp, fltr_df) #-2.525e-05 

mlr_model_forwards <- lm(GPA.Ttl ~ Admission.Grade, fltr_df) #0.005764

mlr_model_forwards <- lm(GPA.Ttl ~ Scholarship, fltr_df) #0.03313

# We select the variable with the largest adjusted R^2 values 
# the next variable will be selected based on the largest adjusted R^2 value when added to the model

```
```{r}

# Function for backward selection using adjusted R^2
backwards_selection_p_value <- function(formula, data, alpha = 0.05) {
  # Fit the full model
  full_model <- lm(formula, data = data)
  
  # Get the p-values of the predictors
  p_values <- summary(full_model)$coefficients[, 4]
  
  # Identify predictors to remove
  predictors_to_remove <- which(p_values > alpha)
  
  # Remove predictors one by one
  for(i in predictors_to_remove) {
    temp_formula <- update(formula,. ~. - terms(formula)[i])
    temp_model <- lm(temp_formula, data = data)
  }
  
  # Return the final model
  return(full_model)
}

mlr_model_backwards <- lm(GPA.Ttl ~ Age + Gender + Admission.Grade + Scholarship, fltr_df)
summary(mlr_model_backwards)

new_model <- backwards_selection_adjusted_R2(GPA.Ttl ~ Age + Gender + Mom.Ed + Mom.Occp + Dad.Ed + Dad.Occp + Admission.Grade + Scholarship, fltr_df)

summary(new_model)
```

